---
title: "DATA550 Midterm Project"
author: "Naziza Bhuiyan, Rou-zhen Chen, Mia Yang, David Kang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
---

```{r, read-cofig, include=FALSE}
WHICH_CONFIG <- Sys.getenv("WHICH_CONFIG")
config_list <- config::get(
    config = WHICH_CONFIG
)
```

```{r, set-chunks-opts, echo=FALSE, eval = TRUE}

knitr::opts_chunk$set(
  echo=FALSE, warning=FALSE, message=FALSE
)
```

```{r, load-packages}

```

## 1. Introduction

### 1.1 Description of Dataset

This dataset comes from a multicenter, double-blind randomized controlled trial evaluating standard F75 versus modified lactose-free, reduced-carbohydrate F75 among hospitalized children with severe acute malnutrition in Kenya and Malawi. It includes detailed baseline demographic, clinical, anthropometric, and biochemical measurements, as well as daily follow-up data through stabilization, discharge, or death. The dataset captures key outcomes such as mortality, diarrhea, appetite, edema, and time to stabilization, allowing for evaluation of treatment effects and clinical predictors of recovery.

### 1.2 Analytical Objectives

The primary analytical objective is to assess whether receiving modified F75 reduces the odds of in-hospital death before stabilization compared to standard F75. Additional objectives include identifying baseline child characteristics—such as diarrhea at admission, MUAC, edema status, age, sex, and HIV status—that independently predict mortality. This analysis will help determine which clinical factors place severely malnourished children at highest risk and whether formula modification provides any protective effect.

### 1.3 Reading in Clean Dataset

```{r, import-dataset, echo=TRUE}
data <- readRDS(
  file = here::here("output", "f75_clean.rds")
) 
```

## 2. Descriptive Statistics

### 2.1 Tabular Analysis

#### Table 1: 
```{r, table-one}
library(dplyr)
library(gtsummary)
library(here)

table1 <- readRDS(
  file = here::here("output/table1.rds")
)

table1

```
At enrollment, a total of 485 children were included in the analysis, with 245 randomized to the Standard F75 arm and 240 to the Modified F75 arm. Baseline demographic and clinical characteristics were comparable across the two treatment groups (Table 1). The median age of participants was 17 months (IQR: 10–26), and the distribution across age tertiles was balanced between arms. Approximately 53% of the cohort was male, with nearly identical sex distributions in both groups. Anthropometric measurements at admission—including MUAC, weight, and height—were similar between treatment arms, indicating comparable nutritional status at baseline. Clinical indicators also demonstrated no meaningful differences: oedema severity, as classified on the 0–3 scale, followed the same pattern across arms; diarrhoea was present in roughly half of participants in each group; and the prevalence of HIV infection was identical at 22–23%. Kwashiorkor, an important marker of edematous malnutrition, was observed in 34% of children overall, with equal proportions in both arms. Collectively, these findings confirm that randomization achieved well-balanced baseline characteristics, minimizing confounding and supporting the validity of subsequent comparisons of clinical outcomes between the Standard and Modified F75 formulations.

## 3. Regression Analysis

### 3.1 Logistic regression for in-hospital mortality before stabilization

We fit a multivariable **logistic regression** model to evaluate whether receiving Modified F75 (intervention) versus Standard F75 was associated with the odds of in-hospital death before stabilization. The binary outcome was death prior to stabilization (1 = died, 0 = survived). The primary exposure was treatment arm (Modified vs Standard F75). The model was adjusted for weight group (≤ median vs > median), diarrhoea, oedema score, age in months, sex, HIV status, MUAC, and presence of kwashiorkor.

#### Table 2: Multivariable logistic regression of survival before stabilization

```{r regression-table}
regression_table_filename <- paste0(
  "regression_table_config_",
  WHICH_CONFIG,
  ".rds"
)

tableregression <- readRDS(
  file = here::here("output", regression_table_filename)
)

tableregression
```

Table 2 presents adjusted odds ratios (ORs) for survival before stabilization from the multivariable logistic regression model for `r config_list$sex`. The primary coefficient of interest is the treatment arm indicator comparing Modified F75 to Standard F75. An OR below 1 for Modified F75 would suggest a reduction in the odds of in-hospital survival prior to stabilization relative to Standard F75, after controlling for baseline clinical severity and demographic covariates. Covariates such as lower MUAC, presence of oedema, diarrhoea, HIV infection, and kwashiorkor are expected to be associated with lower odds of survival, reflecting greater underlying illness severity and malnutrition. The regression framework therefore allows us to estimate the independent association between treatment formulation and mortality while accounting for these potential confounders.

#### Figure 1: Predicted probability of survival by treatment arm

```{r fig-prob-by-arm, fig.cap="Figure 1. Predicted probability of survival before stabilization by treatment arm from the multivariable logistic regression model."}
knitr::include_graphics(
  here::here("output", "logistic_regression_plot.png")
)
```

For each arm, the mean predicted probability of survival (with corresponding uncertainty) is shown, summarizing the combined effects of treatment and covariates in the fitted logistic regression model. A visibly lower predicted probability of survival in the Modified F75 arm compared with the Standard F75 arm does not support the hypothesis that the modified formulation reduces early in-hospital mortality. 


