---
title: "DATA550 Midterm Project"
author: "Naziza Bhuiyan, Rou-zhen Chen, Mia Yang, David Kang"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
  )
```

```{r, load-packages}

```

## 1. Introduction

### 1.1 Description of Dataset

### 1.2 Analytical Objectives

### 1.3 Reading in Clean Dataset

```{r, import-dataset, echo=TRUE}
data <- readRDS(
  file = here::here("output", "f75_clean.rds")
) 
```

## 2. Descriptive Statistics

### 2.1 Tabular Analysis

#### Table 1: 
```{r, table-one}
library(dplyr)
library(gtsummary)
library(here)

table1 <- readRDS(
  file = here::here("output/table1.rds")
)

table1

```
At enrollment, a total of 485 children were included in the analysis, with 245 randomized to the Standard F75 arm and 240 to the Modified F75 arm. Baseline demographic and clinical characteristics were comparable across the two treatment groups (Table 1). The median age of participants was 17 months (IQR: 10–26), and the distribution across age tertiles was balanced between arms. Approximately 53% of the cohort was male, with nearly identical sex distributions in both groups. Anthropometric measurements at admission—including MUAC, weight, and height—were similar between treatment arms, indicating comparable nutritional status at baseline. Clinical indicators also demonstrated no meaningful differences: oedema severity, as classified on the 0–3 scale, followed the same pattern across arms; diarrhoea was present in roughly half of participants in each group; and the prevalence of HIV infection was identical at 22–23%. Kwashiorkor, an important marker of edematous malnutrition, was observed in 34% of children overall, with equal proportions in both arms. Collectively, these findings confirm that randomization achieved well-balanced baseline characteristics, minimizing confounding and supporting the validity of subsequent comparisons of clinical outcomes between the Standard and Modified F75 formulations.

## 3. Regression Analysis

### 3.1 [insert regression type]

#### Table 2:

